Title Goes Here

### 2020-02-24 paper/paper.tex
Describing Core pluging (not gotten to tail recursion transformation yet)

%%
%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,screen]{acmart}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{tcolorbox}

@@

@@

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

@@

>>>>

%%
%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,screen]{acmart}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{tcolorbox}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{stmaryrd}
\usepackage{amsmath}

% \lstset{basicstyle=\ttfamily}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, text width=0.3*\columnwidth, minimum height=1cm, text centered, draw=black, fill=orange!30]
\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

@@

\newcommand{\expr}[1]{\llbracket#1\rrbracket} % Oxford brackets
\newcommand{\rarr}{\rightarrow}
\newcommand{\Rarr}{\Rightarrow}

@@

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Haskell '20]{Haskell '20: Proceedings of the 13th
  ACM SIGPLAN International Symposium on Haskell}{August 27 -- 28, 2020}{New Jersey, United States,}
\acmBooktitle{Haskell '20: ACM SIGPLAN International Symposium on Haskell,
  August 27 -- 28, 2020, New Jersey, United States}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

@@

\section{Core Plugin}

The Core plugin translates marked expressions. Expressions are marked by
the \ttt{externalize} function:

\begin{lstlisting}
externalize :: a -> E a
\end{lstlisting}

So, \ttt{externalize x} marks the expression \ttt{x}.

If an expression already has an EDSL type (a type of the form \ttt{E a} for some
\ttt{a}), then it does the marking procedure ignores it and does not wrap it with
\ttt{externalize}.

\subsection{Overall Control Flow of the Core Plugin}

\begin{tikzpicture}[node distance=2cm]
\node (start) [startstop] {Core from GHC};
  \node (findmark) [process, below of=start]{Find marked expression and erase mark};
    \node (handlecase) [process, below of=findmark]{Mark scrutinee};
    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{Encode as EDSL primitive};
    \node (handleapp) [process, right of=handleprim, xshift=1.6cm]{Wrap with \ttt{Construct} and \ttt{ConstructAp}};

    \node (marksubs1) [process, below of=handleprim]{Mark sub-expressions};
    \node (marksubs2) [process, below of=handleapp]{Mark sub-expressions};

    \node (onebranch) [process, below of=handlecase]{Transform one branch};
    \node (prodmatch) [process, below of=onebranch]{Transform one field in product match of current branch};
    \node (abstractvar) [process, below of=prodmatch]{Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
    \node (markbody) [process, below of=abstractvar]{Mark branch body};
    \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};

  \draw [arrow] (start) -- (findmark);
  \draw [arrow] (findmark) -- node[anchor=east] {Primitive} (handleprim);
  \draw [arrow] (findmark) -- node[anchor=west] {Function application/non-prim var} (handleapp);
  \draw [arrow] (findmark) -- node[anchor=east] {\ttt{case}} (handlecase);

  \draw [arrow] (handleprim) -- (marksubs1);

  \draw [arrow] (handleapp) -- (marksubs2);

  \draw [arrow] (handlecase) -- (onebranch);
  \draw [arrow] (onebranch) -- (prodmatch);
  \draw [arrow] (prodmatch) -- (abstractvar);
  \draw [arrow] (abstractvar) -- (markbody);

  \draw [arrow] (marksubs1) -- (endstep);
  \draw [arrow] (marksubs2) -- (endstep);
  \draw [arrow] (markbody) -- (endstep);

  % \draw (prodmatch) edge[bend right, left] node {While fields remain} (abstractvar);
  % \draw [arrow] (abstractvar.east) to ++(0.5,0) |- node[anchor=west] {While fields remain} (prodmatch.east);
  \draw [arrow] (abstractvar.west) -| ++(-1.5,0) |- node[anchor=south] {While fields remain} (prodmatch.west);

  \draw [arrow] (markbody.west) -| ++(-3,0) |- node[anchor=east, anchor=south] {While branches remain} (onebranch.west);

  \draw [arrow] (endstep.west) -| ++(-8.5,0) |- node[anchor=west, anchor=south] {While marked expressions remain} (findmark.west);

  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);

  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);
  % \draw (onebranch) edge[loop left] node {While branches remain} (onebranch);
  % \Loop[label=\(While\, fields\, remain\), labelstyle={left=5pt}](prodmatch);
  % \path (bodylam) edge [left] (onebranch);
  % \path (bodylam) edge [loop above] node {test} (onebranch);
  % \Loop[label=\(While\, remaining\, branches\, exist\), labelstyle={left=5pt}](onebranch)
  % \draw [arrow] (bodylam) -- (onebranch);
  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
  % \Loop[label=\(While\, branches\, remain\), labelstyle={left=5pt}](onebranch);
  % \Loop[label=\(While\, marked\, expressions\, remain\), labelstyle={left=5pt}](findmark);
\end{tikzpicture}

\clearpage
\subsection{Core Plugin Transformations}

\begin{align*}
  &C\expr{\ttt{externalize (x + y)}} = \ttt{Add (mark x) (mark y)}\\
  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} = \ttt{CaseExp (mark scrutinee) } B\expr{\ttt{\{ ... \}}}\\
  &\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} = \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} = \ttt{OneSumMatchExp }P\expr{\ttt{K x0 ... xN $\rarr$ body}}\\
  &B\expr{\ttt{\{ \}}} = \ttt{EmptyMatch}\\
  &\\
  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} = \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
  &P\expr{\ttt{x $\rarr$ body}} = \ttt{OneProdMatchExp }A\expr{\texttt{$\lambda$ x $\rarr$ mark body}}\\
  &P\expr{\ttt{ $\rarr$ body}} = \ttt{NullaryMatch (mark body)}\\
  &\\
  &A\expr{\ttt{$\lambda$x $\rarr$ body}} = \ttt{Lam (Name uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{   (where \texttt{uniq} is a globally unique identifier)}
\end{align*}
